Safety Analysis of the Molten Salt Fast Reactor Fuel Composition using Moltres

### 2019-03-15 abstract.tex

@@

\begin{abstract}
%
Molten salt reactors (MSR) potentially possess the ability to run for extended periods with minimal shutdown time due to online fuel reprocessing.
Their proposed equilibrium fuel compositions differ from start-up compositions due to burnup of initial fissile material, breeding of new fissile material, and production/removal of fission products.
The impact of changing fuel composition on safety parameters (e.g. reactivity feedback coefficient) must be fully characterized to establish a licensing case for this class of reactors.

Numerous computational tools exist for conventional nuclear reactors.
However, most are not suitable for the study of MSRs.
MSRs differ greatly from conventional solid-fuelled reactors, particularly in their neutronics and thermal-hydraulics behaviours.
They present unique computational challenges that many existing computational tools fail to address effectively.
New MSR simulation tools must capture strong coupling between neutronics and thermal-hydraulics exhibited by delayed neutron precursor movement as well as strong Doppler and density feedback in the fuel salt.
This paper introduces a new simulation tool, Moltres \cite{lindsay_introduction_2018}, for investigating the impact of changing fuel composition on safety parameters.

Moltres is an open source coupled neutronics/thermal hydraulics simulation tool for simulating MSRs.
It is a software package built on the MOOSE finite element framework \cite{gaston_moose:_2009}.
It solves the coupled time-dependent multi-group neutron diffusion, temperature and delayed neutron precursor (DNP) governing equations.
The temperature and DNP equations fully account for fuel advection as the fuel salt flows upwards through the core.

The MSR model studied in this paper is the Molten Salt Fast Reactor (MSFR) \cite{merle-lucotte_launching_2011}.
The MSFR is a reference design for a fast-spectrum molten salt reactor developed under the EVOL and SAMOFAR projects \cite{serp_molten_2014}.

The full paper will present Moltres simulation results of the MSFR reference model with three fuel compositions:
start-up, early life and equilibrium fuel compositions.
The chosen start-up fuel composition is a eutectic mixture of $^{233}$U and $^{232}$Th fluorides in a lithium fluoride molten salt \cite{merle-lucotte_launching_2011}.
This is in line with the main purpose of running the MSFR as a thorium breeder.
Cross-section data for these fuel compositions are generated using Serpent \cite{leppanen_serpent_2015},
which is a continuous-energy Monte Carlo code used in numerous reactor physics applications.
Moltres will solve the coupled neutron transport equation as a steady-state eigenvalue problem to obtain the effective neutron multiplication factor, $k_{eff}$.
For each fuel composition, the temperature distribution and densities will be varied slightly from the steady-state configuration to obtain the change in $k_{eff}$.
Reactivity feedback coefficients will be calculated from these variations and assessed for each fuel compositions.
Finally, the results will be compared with existing data \cite{fiorina_analysis_2012} \cite{fiorina_investigation_2013} of safety parameters for the three fuel compositions.

\end{abstract}

>>>>

\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries

@@

\begin{abstract}
%
        \glspl{MSR} potentially possess the ability to run for extended 
        periods with minimal shutdown time due to online fuel reprocessing.  
        Their proposed equilibrium fuel compositions differ substantially from 
        start-up compositions due to burnup of initial fissile material and 
        breeding of new fissile material, but also fissile material feeds and 
        removal of fission products.  Since the changing fuel composition 
        impacts safety parameters (e.g. reactivity feedback coefficients), a 
        licensing case for this class of reactors  must fully characterize 
        those impacts.

Numerous computational tools exist for conventional nuclear reactors, but 
        \glspl{MSR} present unique computational challenges that many fail to 
        address effectively.  \glspl{MSR} differ profoundly from conventional 
        solid-fuelled reactors, particularly in their neutronics and 
        thermal-hydraulics behaviors.  New \gls{MSR} simulation tools must 
        capture strong coupling between neutronics and thermal-hydraulics 
        exhibited by delayed neutron precursor movement as well as strong 
        Doppler and density feedback in the fuel salt.  This paper investigates 
        the impact of changing fuel composition on safety parameters in the 
        \gls{MSFR}  using a new simulation tool for \glspl{MSR}, Moltres 
        \cite{lindsay_introduction_2018}.

Moltres is an open source coupled neutronics/thermal hydraulics simulation 
        application for simulating \glspl{MSR}. Built on the \gls{MOOSE} finite 
        element framework \cite{gaston_moose:_2009}, Moltres solves the coupled 
        time-dependent multi-group neutron diffusion, temperature, and 
        \gls{DNP} governing equations.  The temperature and DNP equations fully 
        account for fuel advection as the fuel salt flows upwards through the 
        core.

The \gls{MSFR} model studied in this paper
        \cite{merle-lucotte_launching_2011}, is a reference design for a 
        fast-spectrum \gls{MSR} developed under the \gls{EVOL} and 
        \gls{SAMOFAR} projects \cite{serp_molten_2014}.
        The full paper will present results from Moltres simulation of the 
        \gls{MSFR} reference model with three fuel compositions:
        start-up, early life, and equilibrium.

In line with the purpose of the \gls{MSFR} as a thorium breeder, its chosen 
        start-up fuel composition is a eutectic mixture of $^{233}$U and 
        $^{232}$Th fluorides in a lithium fluoride molten salt 
        \cite{merle-lucotte_launching_2011}. We generate group constants for 
        each fuel composition using Serpent \cite{leppanen_serpent_2015}, a 
        continuous-energy Monte Carlo code used in numerous reactor physics 
        applications. For verification, group constants and reactivity 
        coefficients will be compared with existing data 
        \cite{fiorina_analysis_2012,fiorina_investigation_2013} of safety 
        parameters for the three fuel compositions.  
        Using these group constants ($\chi$, $\nu\Sigma_f$, 
        $\Sigma_{g\rightarrow g'}$, $\frac{d\rho}{dT}$, etc.), Moltres then 
        solves for the flux and temperature based on the neutron diffusion 
        equation coupled with navier stokes thermal hydraulics. Transient 
        simulations will establish the spatial distribution of flux, 
        $\phi_g(\vec{r})$ and temperature, $T(\vec{r})$ during transients.
\end{abstract}
